Exploring Local Explanations of Non-linear Models Using Animated Linear Projections

### 2022-01-07 _paper.rmd
typo

---
title: "Exploring Local Explanations of Non-linear Models Using Animated Linear Projections"
shorttitle: Cheem
keywords:
- R
- XAI
- explanable artificial intelligence
- black-box models
- local explanations
- grand tour
- manual tour
- radial tour
author:
#- name: "Nicholas Spyrison"
#   address: |
#     | ORCiD: 0000-0002-8417-0212
#   email: \email{nicholas.spyrison@monash.edu}
#   orcid: 0000-0002-8417-0212
#- name: "Dianne Cook"
#   affiliation: Monash University
#   address: |
#     | ORCiD: 000-0002-3813-7155
#   orcid: 000-0002-3813-7155
#- name: "Kimbal Marriott"
#   affiliation: Monash University
#   address: |
#     | ORCiD: 0000-0002-9813-0377
#   orcid: 0000-0002-9813-0377
abstract: |
 Artificial Intelligence (AI) has seen a revitalization in recent years with the prevalence of increasingly hard-to-interpret black-box models. The increased predictive power comes at the cost of interpretability, which has led to the emergence of eXplainable AI (XAI). XAI attempts to shed light on how models are using predictors, to arrive at a prediction, with a point estimate of the linear variable importance in the vicinity of each observation. These are simply linear projections, and as such can be further explored interactively to better understand the interaction between variables used to make predictions, across the predictive model surface. Here we describe interactive linear interpolation used for exploration at any observation, and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (footbal salaries, house prices) response variables. The methods are implemented in the __R__ package __cheem__, available on CRAN.
# documentclass: jdssv
#classoption: article
output:
  ## bookdown::pdf_document2: allows \@ref()
  bookdown::pdf_document2:
  ## pdf_document: requires \ref{} 
  #pdf_document:
    keep_tex: true
    toc: false
    #template: template/jdssv_template.tex
    number_sections: yes
    fig_caption: yes
bibliography: cheem_paper.bib
#appendix:
# title:
#   formatted: "Methods for understanding the variable importance of local explanations of black-box models"
#   plain:     "Methods for understanding the variable importance of local explanations of black-box models"
#   short:     "cheem: radial tour scrutinizing local explanations"
editor_options:
  chunk_output_type: console
urlcolor: blue
linkcolor: red
header-includes:
 - \usepackage{setspace} ## for title page spacing
 - \usepackage{hyperref} ## for all sorts of linking
---

>>>>

---
title: "Exploring Local Explanations of Non-linear Models Using Animated Linear Projections"
shorttitle: Cheem
keywords:
- R
- XAI
- explanable artificial intelligence
- black-box models
- local explanations
- grand tour
- manual tour
- radial tour
author:
#- name: "Nicholas Spyrison"
#   address: |
#     | ORCiD: 0000-0002-8417-0212
#   email: \email{nicholas.spyrison@monash.edu}
#   orcid: 0000-0002-8417-0212
#- name: "Dianne Cook"
#   affiliation: Monash University
#   address: |
#     | ORCiD: 000-0002-3813-7155
#   orcid: 000-0002-3813-7155
#- name: "Kimbal Marriott"
#   affiliation: Monash University
#   address: |
#     | ORCiD: 0000-0002-9813-0377
#   orcid: 0000-0002-9813-0377
abstract: |
 Artificial Intelligence (AI) has seen a revitalization in recent years with the prevalence of increasingly hard-to-interpret black-box models. The increased predictive power comes at the cost of interpretability, which has led to the emergence of eXplainable AI (XAI). XAI attempts to shed light on how models are using predictors, to arrive at a prediction, with a point estimate of the linear variable importance in the vicinity of each observation. These are simply linear projections, and as such can be further explored interactively to better understand the interaction between variables used to make predictions, across the predictive model surface. Here we describe interactive linear interpolation used for exploration at any observation, and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (football salaries, house prices) response variables. The methods are implemented in the __R__ package __cheem__, available on CRAN.
# documentclass: jdssv
#classoption: article
output:
  ## bookdown::pdf_document2: allows \@ref()
  bookdown::pdf_document2:
  ## pdf_document: requires \ref{} 
  #pdf_document:
    keep_tex: true
    toc: false
    #template: template/jdssv_template.tex
    number_sections: yes
    fig_caption: yes
bibliography: cheem_paper.bib
#appendix:
# title:
#   formatted: "Methods for understanding the variable importance of local explanations of black-box models"
#   plain:     "Methods for understanding the variable importance of local explanations of black-box models"
#   short:     "cheem: radial tour scrutinizing local explanations"
editor_options:
  chunk_output_type: console
urlcolor: blue
linkcolor: red
header-includes:
 - \usepackage{setspace} ## for title page spacing
 - \usepackage{hyperref} ## for all sorts of linking
---

### 2022-01-07 _paper.tex
typo

\begin{document}
\maketitle
\begin{abstract}
Artificial Intelligence (AI) has seen a revitalization in recent years with the prevalence of increasingly hard-to-interpret black-box models. The increased predictive power comes at the cost of interpretability, which has led to the emergence of eXplainable AI (XAI). XAI attempts to shed light on how models are using predictors, to arrive at a prediction, with a point estimate of the linear variable importance in the vicinity of each observation. These are simply linear projections, and as such can be further explored interactively to better understand the interaction between variables used to make predictions, across the predictive model surface. Here we describe interactive linear interpolation used for exploration at any observation, and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (footbal salaries, house prices) response variables. The methods are implemented in the \textbf{R} package \textbf{cheem}, available on CRAN.
\end{abstract}

>>>>

\begin{document}
\maketitle
\begin{abstract}
Artificial Intelligence (AI) has seen a revitalization in recent years with the prevalence of increasingly hard-to-interpret black-box models. The increased predictive power comes at the cost of interpretability, which has led to the emergence of eXplainable AI (XAI). XAI attempts to shed light on how models are using predictors, to arrive at a prediction, with a point estimate of the linear variable importance in the vicinity of each observation. These are simply linear projections, and as such can be further explored interactively to better understand the interaction between variables used to make predictions, across the predictive model surface. Here we describe interactive linear interpolation used for exploration at any observation, and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (football salaries, house prices) response variables. The methods are implemented in the \textbf{R} package \textbf{cheem}, available on CRAN.
\end{abstract}
